Reddouane BERRAMDANE
Independent Researcher
reddoma@gmail.com
[City, Country]
[Phone number]

July 8, 2026

Dear Editor,

I am pleased to submit our manuscript entitled "Discrete-to-Continuum 
Energy Matching in Solid-State Defect Centers: Statistical Raman 
Screening of Atomic and Defect Spectra Near the NV Zero-Field Splitting" 
for consideration for publication in [Journal Name].

This paper presents a statistical screening framework that identifies 
atomic spectra containing energy-level pairs near the nitrogen-vacancy 
(NV) center zero-field splitting in diamond. The work is the third 
installment in the Discrete-to-Continuum Quantum Energy Matching (DTQEM) 
series and provides quantitative support for the hypothesis that 
solid-state defect centers require an external phonon continuum for 
Raman processes, while dense atomic spectra may possess such 
flexibility internally.

Key findings include:
• Fe I exhibits 69 true Raman pairs (q = 0.0012 after FDR), surviving 
  the most conservative null model.
• Defect spectra (NV, SiV, GeV) possess zero Raman pairs, confirming 
  their structural inability to supply internal energy matching.
• The study demonstrates the critical importance of structure-preserving 
  null models in spectral screening.

We believe this work is suitable for [Journal Name] because it addresses 
fundamental questions in quantum sensing and defect-center physics using 
a reproducible, statistically rigorous methodology. The framework is 
target-agnostic and can be applied broadly to spectral databases.

This manuscript is original, has not been published previously, and is 
not under consideration elsewhere. All authors have approved the 
manuscript and agree with its submission to [Journal Name].

We suggest the following potential referees:
[To be added based on journal requirements]

Thank you for considering our manuscript. We look forward to your 
response.

Sincerely,

Reddouane BERRAMDANE
Independent Researcher
